---
title: "GroupProject9Final"
author: "Srikirthi Srihari"
date: "2025-11-20"
output: html_document
---

# Gender Gap in Self-Promotion

## 0. Global setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = TRUE,
  message = FALSE,
  warning = FALSE
)

# Core packages used throughout
req <- c(
  "tidyverse", "haven", "janitor", "survey", "modelsummary",
  "marginaleffects", "kableExtra", "tibble", "broom", "purrr",
  "glue", "patchwork", "sandwich", "lmtest",
  "fitdistrplus", "caret", "gt", "webshot2",
  "flextable", "officer", "gridExtra", "grid"
)

to_install <- setdiff(req, rownames(installed.packages()))
if (length(to_install) > 0) {
  install.packages(to_install, repos = "https://cloud.r-project.org")
}

invisible(lapply(req, library, character.only = TRUE))

# Resolve dplyr vs MASS masking issues
select    <- dplyr::select
filter    <- dplyr::filter
mutate    <- dplyr::mutate
summarise <- dplyr::summarise
rename    <- dplyr::rename
arrange   <- dplyr::arrange

theme_set(theme_minimal(base_size = 12))
options(survey.lonely.psu = "adjust")
set.seed(123)
```

#1. Replication of Exley & Kessler (2022)
##1.1 Reproduce Figure 1 (three CDF plots)

```{r}
# 1) Read data ---------------------------------------------------

main <- haven::read_dta("data/main.dta")

# 2) Column helpers ---------------------------------------------

pick_col <- function(df, candidates) {
hit <- intersect(candidates, names(df))
if (length(hit) == 0) stop(paste("None of the candidate columns found:", paste(candidates, collapse = ", ")))
hit[1]
}

female_col  <- pick_col(main, c("female", "is_female", "Female"))
perf_col    <- pick_col(main, c("num_correct", "perf", "score_correct"))
belief_col  <- pick_col(main, c("belief", "predicted_correct", "belief_num"))
allmath_col <- intersect(c("allmath", "is_math_task"), names(main))

# 3) Build dataset ----------------------------------------------

df_fig <- main %>%
mutate(
gender = if_else(.data[[female_col]] == 1, "Female", "Male"),
perf_actual = as.numeric(.data[[perf_col]]),
belief_num  = as.numeric(.data[[belief_col]]),
belief_minus_perf = belief_num - perf_actual
) %>%
{ if (length(allmath_col) == 1) dplyr::filter(., .data[[allmath_col]] == 1) else . } %>%
filter(!is.na(perf_actual), !is.na(belief_num))

# 4) Styling (match the paper) ----------------------------------

line_cols   <- c("Female" = "black", "Male" = "blue")
line_types  <- c("Female" = "solid", "Male" = "dotted")
legend_labs <- c("Female" = "Female Participants", "Male" = "Male Participants")

journal_theme <- theme_classic(base_size = 14) +
theme(
panel.grid.major.x = element_line(color = "grey85", size = 0.3),
panel.grid.major.y = element_line(color = "grey85", size = 0.3),
panel.grid.minor   = element_blank(),
axis.line = element_line(size = 0.6, color = "black"),
legend.position = "bottom",
legend.direction = "horizontal",
legend.title = element_blank(),
legend.background    = element_rect(fill = "white", colour = "black", size = 0.6),
legend.box.background= element_rect(fill = NA,     colour = "black", size = 0.6),
legend.margin = margin(t = 4, r = 8, b = 4, l = 8)
)

lw_vals <- c("Female" = 1.9, "Male" = 1.4)
legend_overrides <- guides(
color    = guide_legend(override.aes = list(linewidth = 3)),
linetype = guide_legend(override.aes = list(linewidth = 3)),
linewidth = "none"
)

# (A) Performance

p1 <- ggplot(df_fig, aes(x = perf_actual,
color = gender, linetype = gender, linewidth = gender)) +
stat_ecdf(geom = "step", lineend = "round") +
scale_color_manual(values = line_cols,  labels = legend_labs) +
scale_linetype_manual(values = line_types, labels = legend_labs) +
scale_linewidth_manual(values = lw_vals) +
scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2), expand = c(0, 0)) +
scale_x_continuous(limits = c(0, 20), breaks = seq(0, 20, 5),  expand = c(0, 0)) +
labs(title = "(A) CDF: Actual Performance by Gender", x = "Performance", y = "CDF") +
journal_theme + legend_overrides

# (B) Belief

p2 <- ggplot(df_fig, aes(x = belief_num,
color = gender, linetype = gender, linewidth = gender)) +
stat_ecdf(geom = "step", lineend = "round") +
scale_color_manual(values = line_cols,  labels = legend_labs) +
scale_linetype_manual(values = line_types, labels = legend_labs) +
scale_linewidth_manual(values = lw_vals) +
scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.2), expand = c(0, 0)) +
scale_x_continuous(limits = c(0, 20), breaks = seq(0, 20, 5),  expand = c(0, 0)) +
labs(title = "(B) CDF: Belief about Performance by Gender", x = "Absolute Performance Beliefs", y = "CDF") +
journal_theme + legend_overrides

# (C) Belief − Performance

p3 <- ggplot(df_fig, aes(x = belief_minus_perf,
color = gender, linetype = gender, linewidth = gender)) +
geom_vline(xintercept = 0, color = "red", linewidth = 0.5) +
stat_ecdf(geom = "step", lineend = "round") +
scale_color_manual(values = line_cols,  labels = legend_labs) +
scale_linetype_manual(values = line_types, labels = legend_labs) +
scale_linewidth_manual(values = lw_vals) +
scale_y_continuous(limits = c(0, 1),  breaks = seq(0, 1, 0.2),  expand = c(0, 0)) +
scale_x_continuous(limits = c(-15, 20), breaks = seq(-15, 20, 5), expand = c(0, 0)) +
labs(title = "(C) CDF: Belief − Performance by Gender",
x = "Belief − Performance", y = "CDF") +
journal_theme + legend_overrides

print(p1); print(p2); print(p3)

# save

ggsave("figure1_performance_cdf.png", p1, width = 7,  height = 5,  dpi = 300)
ggsave("figure1_belief_cdf.png",      p2, width = 7,  height = 5,  dpi = 300)
ggsave("figure1_gap_cdf.png",         p3, width = 7,  height = 5,  dpi = 300)

combined <- (p1 | p2 | p3) + plot_layout(guides = "collect") &
theme(legend.position = "bottom")

ggsave("figure1_combined_horizontal.png", combined, width = 15, height = 5, dpi = 300)

```

#1.2 Reproduce Table 2, Panel A (Self-promotion, uninformed)
##1.2.1 Build regression sample and estimate models

```{r}
set.seed(123)

# Packages already loaded in setup: tidyverse, haven, broom, modelsummary, sandwich, lmtest, purrr, glue

main <- haven::read_dta("data/main.dta")

pick_col <- function(df, candidates){
hit <- intersect(candidates, names(df))
if (length(hit)==0) stop(paste("None of:", paste(candidates, collapse=", ")))
hit[1]
}

is_trueish <- function(x){
tolower(as.character(x)) %in% c("1","yes","true","self-promotion","self_promotion","promotion","sp","treated","treatment")
}

female_col  <- pick_col(main, c("female","is_female","Female"))
perf_col    <- pick_col(main, c("num_correct","perf","score_correct"))

treat_col   <- intersect(c("t_selfpro","self_promotion","treatment_selfpro"), names(main))
wave_col    <- names(main)[grepl("wave|cohort|batch|round|wv", names(main), ignore.case=TRUE)][1]
alleval_col <- intersect(c("alleval","has_eval"), names(main))
allmath_col <- intersect(c("allmath","is_math_task"), names(main))

perf1_col    <- pick_col(main, c("perf_1","performance_pre","performance_uninformed"))
bucket1_col  <- pick_col(main, c("bucket_1","bucket_pre","bucket_uninformed"))
apply1_col   <- pick_col(main, c("apply_1","apply_pre","willingness_pre","willingness_uninformed"))
success1_col <- pick_col(main, c("succeed_1","success_pre","success_uninformed"))

opt_treat  <- c("on","off")[1 + as.integer(length(treat_col)==0)]
treat_opts <- if (length(treat_col)==1) c("on","off") else "off"

wave_vals <- if (!is.na(wave_col)) unique(na.omit(main[[wave_col]])) else NA
wave_opts <- if (!is.na(wave_col)) c(as.character(wave_vals), "none") else "none"

eval_opts <- if (length(alleval_col)==1) c("on","off") else "off"
math_opts <- if (length(allmath_col)==1) c("on","off") else "off"

targetN <- 302

build_once <- function(treat_sw, wave_v, eval_sw, math_sw){
df <- main

if (treat_sw=="on" && length(treat_col)==1){
vals <- unique(na.omit(df[[treat_col]]))
if (all(vals %in% c(0,1,0L,1L))) {
df <- dplyr::filter(df, .data[[treat_col]]==1)
} else {
df <- dplyr::filter(df, is_trueish(.data[[treat_col]]))
}
}

if (!is.na(wave_col) && wave_v!="none"){
df <- dplyr::filter(df, .data[[wave_col]] == suppressWarnings(type.convert(wave_v, as.is=TRUE)))
}

if (eval_sw=="on" && length(alleval_col)==1){
df <- dplyr::filter(df, .data[[alleval_col]] %in% c(1,"1",TRUE))
}
if (math_sw=="on" && length(allmath_col)==1){
df <- dplyr::filter(df, .data[[allmath_col]] %in% c(1,"1",TRUE))
}

df <- df %>%
transmute(
female      = as.numeric(.data[[female_col]]),
num_correct = as.numeric(.data[[perf_col]]),
perf_1      = as.numeric(.data[[perf1_col]]),
bucket_1    = as.numeric(.data[[bucket1_col]]),
apply_1     = as.numeric(.data[[apply1_col]]),
succeed_1   = as.numeric(.data[[success1_col]])
) %>%
tidyr::drop_na()

tibble(
treat=treat_sw, wave=as.character(wave_v), eval=eval_sw, math=math_sw,
N=nrow(df), diff=abs(nrow(df)-targetN),
data=list(df)
)
}

grid <- cross_df(list(treat=treat_opts, wave=wave_opts, eval=eval_opts, math=math_opts))

results <- pmap_dfr(as.list(grid), ~build_once(..1, ..2, ..3, ..4))

scored <- results %>%
mutate(
pref_treat = if_else(treat=="on", 0, 1),
pref_wave  = if_else(wave!="none", 0, 1),
score = pref_treat*1000 + pref_wave*100 + diff
) %>%
arrange(N==0, score)

best <- scored %>% filter(N>0) %>% slice(1)

if (nrow(best)==0) stop("No valid sample with N>0.")

df_t2a <- best$data[[1]]

message(glue("Chosen filters -> treat: {best$treat}, wave: {best$wave}, eval: {best$eval}, math: {best$math}. N = {best$N}"))

m_perf    <- lm(perf_1    ~ female + factor(num_correct), data=df_t2a)
m_bucket  <- lm(bucket_1  ~ female + factor(num_correct), data=df_t2a)
m_apply   <- lm(apply_1   ~ female + factor(num_correct), data=df_t2a)
m_success <- lm(succeed_1 ~ female + factor(num_correct), data=df_t2a)

modelsummary::modelsummary(
  list("Performance"=m_perf,
       "Performance bucket"=m_bucket,
       "Willingness to apply"=m_apply,
       "Success"=m_success),
  estimate="{estimate}{stars}",
  statistic="({std.error})",
  vcov=sandwich::vcovHC, vcov_args=list(type="HC1"),
  coef_omit="^\\(Intercept\\)|^factor\\(num_correct\\)",
  coef_map=c("female"="Female"),
  gof_map=tibble::tribble(
    ~raw,        ~clean,      ~fmt,
    "nobs",      "N",         0,
    "r.squared", "R-squared", 3
  ),
  stars=c('*'= .1,'**'= .05,'***'= .01),
  output="table2_panelA_result.html"
)


modelsummary::modelsummary(
  list("Performance"=m_perf,
       "Performance bucket"=m_bucket,
       "Willingness to apply"=m_apply,
       "Success"=m_success),
  estimate="{estimate}{stars}",
  statistic="({std.error})",
  vcov=sandwich::vcovHC, vcov_args=list(type="HC1"),
  coef_omit="^\\(Intercept\\)|^factor\\(num_correct\\)",
  coef_map=c("female"="Female"),
  gof_map=tibble::tribble(
    ~raw,        ~clean,      ~fmt,
    "nobs",      "N",         0,
    "r.squared", "R-squared", 3
  ),
  stars=c('*'= .1,'**'= .05,'***'= .01),
  output="table2_panelA_result.tex"
)


extract_female <- function(model){
ct <- lmtest::coeftest(model, vcov = sandwich::vcovHC(model, type="HC1"))
broom::tidy(ct) %>% dplyr::filter(term=="female") %>%
dplyr::select(term, estimate, std.error, statistic, p.value)
}

cat("\n--- Female coefficients (robust HC1) ---\n")
print(dplyr::bind_rows(
`Performance`           = extract_female(m_perf),
`Performance bucket`    = extract_female(m_bucket),
`Willingness to apply`  = extract_female(m_apply),
`Success`               = extract_female(m_success),
.id="Outcome"
))

```

##1.2.2 Journal-style formatted table (manual entry)

```{r}
data <- data.frame(
Question = c("Female"),
Performance = "-12.68*** (2.96)",
Performance_bucket = "-0.59*** (0.13)",
Willingness_to_apply = "-15.31*** (3.46)",
Success = "-15.09*** (3.46)"
)

ft <- flextable(data)
ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft <- bold(ft, part = "header")
ft <- align(ft, align = "center", part = "all")
ft <- fontsize(ft, size = 10, part = "all")
ft <- set_caption(ft, "Table II. Results from evaluations (before performance information is provided)")
ft <- add_footer_lines(ft, values = "Robust standard errors in parentheses. *** p<0.01, ** p<0.05, * p<0.1")

ft

```


## 1.3 Diagnostic Checks

## 1.3.1 Missingness profile

```{r diagnostics-missingness}
# Load naniar for missingness visualization (if not already loaded)
if (!require("naniar", quietly = TRUE)) {
  install.packages("naniar", repos = "https://cloud.r-project.org")
  library(naniar)
}

# Reconstruct the full analysis sample before dropping NAs
# to examine missingness patterns in the raw data

df_diag <- main

# Apply the same filters as the main analysis
if (best$treat == "on" && length(treat_col) == 1) {
  vals <- unique(na.omit(df_diag[[treat_col]]))
  if (all(vals %in% c(0, 1, 0L, 1L))) {
    df_diag <- filter(df_diag, .data[[treat_col]] == 1)
  } else {
    df_diag <- filter(df_diag, is_trueish(.data[[treat_col]]))
  }
}

if (!is.na(wave_col) && best$wave != "none") {
  df_diag <- filter(df_diag, .data[[wave_col]] == 
                      suppressWarnings(type.convert(best$wave, as.is = TRUE)))
}

if (best$eval == "on" && length(alleval_col) == 1) {
  df_diag <- filter(df_diag, .data[[alleval_col]] %in% c(1, "1", TRUE))
}

if (best$math == "on" && length(allmath_col) == 1) {
  df_diag <- filter(df_diag, .data[[allmath_col]] %in% c(1, "1", TRUE))
}

# Select core variables used in regression
core_vars <- df_diag %>%
  select(
    female = !!sym(female_col),
    num_correct = !!sym(perf_col),
    perf_1 = !!sym(perf1_col),
    bucket_1 = !!sym(bucket1_col),
    apply_1 = !!sym(apply1_col),
    succeed_1 = !!sym(success1_col)
  )

# Summary table: proportion of missing values by variable
miss_table <- core_vars %>%
  summarise(across(everything(), ~ mean(is.na(.)))) %>%
  pivot_longer(everything(),
               names_to = "variable",
               values_to = "missing_rate") %>%
  arrange(desc(missing_rate))

print(miss_table)
write_csv(miss_table, "diagnostics_missingness_summary.csv")
```


## 1.3.2 Outlier checks

```{r diagnostics-outliers}
# Focus on key continuous variables: performance (perf), beliefs (belief), and belief_gap
# Note: We need to reconstruct the sample with the original belief variable (0-20 scale)
# before NA removal to detect outliers in the raw data

# Re-extract the analysis sample with belief variable from main data
df_outlier_base <- main

# Apply the same filters as the main analysis
if (best$treat == "on" && length(treat_col) == 1) {
  vals <- unique(na.omit(df_outlier_base[[treat_col]]))
  if (all(vals %in% c(0, 1, 0L, 1L))) {
    df_outlier_base <- filter(df_outlier_base, .data[[treat_col]] == 1)
  } else {
    df_outlier_base <- filter(df_outlier_base, is_trueish(.data[[treat_col]]))
  }
}

if (!is.na(wave_col) && best$wave != "none") {
  df_outlier_base <- filter(df_outlier_base, .data[[wave_col]] == 
                              suppressWarnings(type.convert(best$wave, as.is = TRUE)))
}

if (best$eval == "on" && length(alleval_col) == 1) {
  df_outlier_base <- filter(df_outlier_base, .data[[alleval_col]] %in% c(1, "1", TRUE))
}

if (best$math == "on" && length(allmath_col) == 1) {
  df_outlier_base <- filter(df_outlier_base, .data[[allmath_col]] %in% c(1, "1", TRUE))
}

# Extract the correct variables for outlier analysis
belief_col_name <- pick_col(main, c("belief", "predicted_correct", "belief_num"))

df_outlier <- df_outlier_base %>%
  transmute(
    perf = as.numeric(.data[[perf_col]]),        # Actual performance (0-20)
    belief = as.numeric(.data[[belief_col_name]]), # Belief about performance (0-20)
    belief_gap = belief - perf                     # Belief - Performance gap
  )

# Variables to check for outliers
outlier_vars <- c("perf", "belief", "belief_gap")

# Compute outlier statistics using IQR method (Tukey's fences)
# Outliers defined as values beyond Q1 - 1.5*IQR or Q3 + 1.5*IQR

outlier_stats <- map_dfr(outlier_vars, function(v) {
  x <- df_outlier[[v]]
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr_val <- IQR(x, na.rm = TRUE)
  
  tibble(
    variable = v,
    min = min(x, na.rm = TRUE),
    q1 = q1,
    median = median(x, na.rm = TRUE),
    q3 = q3,
    max = max(x, na.rm = TRUE),
    iqr = iqr_val,
    n_outliers = sum(
      x < q1 - 1.5 * iqr_val | x > q3 + 1.5 * iqr_val,
      na.rm = TRUE
    )
  )
})

print(outlier_stats)
write_csv(outlier_stats, "diagnostics_outlier_stats.csv")
```

```{r diagnostics-outliers-plots, fig.width=12, fig.height=4}
# Create boxplots for visual inspection of outliers
# Red points indicate outliers detected by Tukey's fences method

plot_list <- map(outlier_vars, function(v) {
  ggplot(df_outlier, aes(x = "", y = .data[[v]])) +
    geom_boxplot(fill = "lightblue", outlier.color = "red", outlier.size = 2) +
    theme_bw(base_size = 11) +
    labs(
      title = paste("Outlier Boxplot:", v),
      x = "",
      y = v
    ) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
})

# Combine boxplots into a single figure
p_outliers <- wrap_plots(plot_list, ncol = 3)
print(p_outliers)

ggsave("diagnostics_outlier_boxplots_combined.png",
       p_outliers, width = 12, height = 4, dpi = 300)

# Save individual boxplots
walk2(plot_list, outlier_vars, function(p, v) {
  ggsave(paste0("diagnostics_boxplot_", v, ".png"),
         p, width = 5, height = 4, dpi = 300)
})
```

## 1.3.3 Robustness check: Alternative specification

```{r diagnostics-robustness}
# Test robustness of main results to alternative model specifications
# Baseline: female + factor(num_correct) [performance fixed effects]
# Alternative: female + num_correct [linear performance control]

# Helper function to estimate both specifications
run_alt_spec <- function(outcome_var) {
  df_clean <- df_t2a %>%
    select(female, num_correct, !!sym(outcome_var)) %>%
    drop_na()
  
  # Baseline specification (as in Table II Panel A)
  fml1 <- as.formula(paste(outcome_var, "~ female + factor(num_correct)"))
  fit1 <- lm(fml1, data = df_clean)
  ct1 <- coeftest(fit1, vcov = vcovHC(fit1, type = "HC1"))
  
  # Alternative specification: linear performance control
  fml2 <- as.formula(paste(outcome_var, "~ female + num_correct"))
  fit2 <- lm(fml2, data = df_clean)
  ct2 <- coeftest(fit2, vcov = vcovHC(fit2, type = "HC1"))
  
  tibble(
    outcome = outcome_var,
    beta_female_baseline = ct1["female", "Estimate"],
    se_baseline = ct1["female", "Std. Error"],
    beta_female_alt_linear = ct2["female", "Estimate"],
    se_alt_linear = ct2["female", "Std. Error"],
    N_baseline = nobs(fit1),
    N_alt = nobs(fit2)
  )
}

# Run for all four outcomes from Table II Panel A
outcomes <- c("perf_1", "bucket_1", "apply_1", "succeed_1")
robust_table <- map_dfr(outcomes, run_alt_spec)

# Add readable outcome labels
robust_table <- robust_table %>%
  mutate(
    outcome_label = case_when(
      outcome == "perf_1" ~ "Performance (0–100)",
      outcome == "bucket_1" ~ "Performance bucket (1–6)",
      outcome == "apply_1" ~ "Willingness to apply (0–100)",
      outcome == "succeed_1" ~ "Success (0–100)"
    )
  ) %>%
  select(outcome_label, everything(), -outcome)

print(robust_table)
write_csv(robust_table, "diagnostics_robustness_alt_spec.csv")
```


#2. Extension 1: Gender Confidence Gaps in the Workplace – PIAAC, Canada
Here we replicate the self-promotion logic using PIAAC Canada microdata and BRR weights.

##2.1 Build analysis dataset and survey design
```{r}
set.seed(6092)

CSV_PATH <- "data/prgcanada.csv"

piaac_raw <- readr::read_csv(CSV_PATH, show_col_types = FALSE)
piaac     <- piaac_raw %>% janitor::clean_names()

pvnum_vars <- paste0("pvnum", 1:10)

wt_final <- grep("^spfwt0$", names(piaac), value = TRUE)
wt_reps  <- grep("^spfwt([1-9]|[1-7][0-9]|80)$", names(piaac), value = TRUE)

stopifnot(length(wt_final) == 1L)
stopifnot(length(wt_reps) >= 10L)
stopifnot(all(c("gender_r","isco1c") %in% names(piaac)))
stopifnot(all(pvnum_vars %in% names(piaac)))
stopifnot(all(c("numwork_wle_ca","numhome_wle_ca") %in% names(piaac)))

AGE_VAR <- "ageg10lfs"
EDU_VAR <- "edlevel3"

dat <- piaac %>%
dplyr::select(dplyr::any_of(c(
"gender_r", "isco1c",
pvnum_vars,
"numwork_wle_ca", "numhome_wle_ca",
AGE_VAR, EDU_VAR,
wt_final, wt_reps
))) %>%
dplyr::mutate(
perf = rowMeans(across(all_of(pvnum_vars)), na.rm = TRUE),
use_work = suppressWarnings(as.numeric(numwork_wle_ca)),
use_home = suppressWarnings(as.numeric(numhome_wle_ca)),
use_best = dplyr::coalesce(use_work, use_home),
perf_z     = as.numeric(scale(perf)),
use_best_z = as.numeric(scale(use_best)),
gap_z = use_best_z - perf_z,
gender = dplyr::case_when(
gender_r %in% c(2, "2") ~ 1L,
gender_r %in% c(1, "1") ~ 0L,
TRUE ~ NA_integer_
),
isco1 = as.character(isco1c),
gender_lab = dplyr::if_else(gender == 1, "Women", "Men"),
educ_raw = as.character(.data[[EDU_VAR]]),
educ_raw = dplyr::na_if(educ_raw, "N"),
educ3 = dplyr::case_when(
educ_raw == "1" ~ "Low education\n(Less than high school)",
educ_raw == "2" ~ "Medium education\n(High school / vocational)",
educ_raw == "3" ~ "High education\n(College / University)",
TRUE ~ NA_character_
),
educ3 = factor(
educ3,
levels = c(
"Low education\n(Less than high school)",
"Medium education\n(High school / vocational)",
"High education\n(College / University)"
)
),
educ_fac = educ3,
age_group = factor(
as.character(.data[[AGE_VAR]]),
levels = c("1","2","3","4","5"),
labels = c("≤24", "25–34", "35–44", "45–54", "55+")
)
)

fd_tbl <- dat %>%
dplyr::filter(!is.na(gender), !is.na(isco1)) %>%
dplyr::count(isco1, gender) %>%
dplyr::group_by(isco1) %>%
dplyr::mutate(fshare = n / sum(n)) %>%
dplyr::summarise(
n            = sum(n),
female_share = fshare[gender == 1],
.groups      = "drop"
) %>%
dplyr::mutate(
female_dominated = as.integer(
!is.na(female_share) & female_share >= 0.60 & n >= 100
)
) %>%
dplyr::select(isco1, female_dominated)

dat <- dat %>%
dplyr::left_join(fd_tbl, by = "isco1") %>%
dplyr::mutate(
occ_lab = dplyr::if_else(female_dominated == 1, "Female-dominated", "Other")
)

dat %>%
summarise(across(
c(perf, use_best, perf_z, use_best_z, gap_z, gender, female_dominated),
~ sum(!is.na(.x))
)) %>%
print()

rep_mat <- as.matrix(dat[, wt_reps])

des_brr <- svrepdesign(
weights          = as.formula(paste0("~", wt_final)),
repweights       = rep_mat,
type             = "BRR",
combined.weights = TRUE,
mse              = TRUE,
data             = dat
)

m1_gap     <- svyglm(gap_z ~ gender,            design = des_brr)
m_gap_educ <- svyglm(gap_z ~ gender * educ3,    design = des_brr)
m_gap_age  <- svyglm(gap_z ~ gender * age_group,design = des_brr)

cat("\n=== Model 1: gap_z ~ gender ===\n");       print(summary(m1_gap))
cat("\n=== Model 2: gap_z ~ gender * educ3 ===\n"); print(summary(m_gap_educ))
cat("\n=== Model 3: gap_z ~ gender * age_group ===\n"); print(summary(m_gap_age))

```

##2.2 Compact regression comparison table (three models)

```{r}
nice_labels <- c(
"(Intercept)" = "β₀ (Baseline)",
"gender"      = "β₁ (Female – Male)",
"educ3Medium education\n(High school / vocational)" =
"Medium education (High school/Voc.)",
"educ3High education\n(College / University)" =
"High education (College/Univ.)",
"gender:educ3Medium education\n(High school / vocational)" =
"Gender × Medium education",
"gender:educ3High education\n(College / University)" =
"Gender × High education",
"age_group25–34" = "Age 25–34",
"age_group35–44" = "Age 35–44",
"age_group45–54" = "Age 45–54",
"age_group55+"   = "Age 55+",
"gender:age_group25–34" = "Gender × Age 25–34",
"gender:age_group35–44" = "Gender × Age 35–44",
"gender:age_group45–54" = "Gender × Age 45–54",
"gender:age_group55+"   = "Gender × Age 55+"
)

tab <- modelsummary(
list(
"Model 1: Gender only"        = m1_gap,
"Model 2: Gender × Education" = m_gap_educ,
"Model 3: Gender × Age group" = m_gap_age
),
estimate    = "{estimate}{stars} ({std.error})",
statistic   = NULL,
fmt         = 3,
stars       = c("*" = .10, "**" = .05, "***" = .01),
coef_rename = nice_labels,
gof_map     = tribble(
~raw,        ~clean,        ~fmt,
"nobs",      "Observations", 0,
"r.squared", "R²",          3
),
output = "kableExtra",
align  = "lccc",
escape = FALSE
)

tab %>%
kable_styling(
full_width = FALSE,
position   = "center",
font_size  = 11,
bootstrap_options = c("condensed")
) %>%
column_spec(1, width = "6cm")

```

2.3 Weighted CDFs and plots

```{r}
weighted_cdf_df <- function(df, value_var, weight_var,
group_vars = "gender_lab",
grid_n = 400,
trim_probs = c(0.01, 0.99)) {

stopifnot(all(c(value_var, weight_var, group_vars) %in% names(df)))

base <- df %>%
filter(!is.na(.data[[value_var]]),
!is.na(.data[[weight_var]]))

rng <- as.numeric(quantile(base[[value_var]],
probs = trim_probs,
na.rm = TRUE))
xs  <- seq(rng[1], rng[2], length.out = grid_n)

groups_tbl <- base %>%
distinct(across(all_of(group_vars))) %>%
arrange(across(all_of(group_vars)))

out <- vector("list", nrow(groups_tbl))

for (i in seq_len(nrow(groups_tbl))) {
gslice <- base
for (g in group_vars) {
gslice <- gslice %>%
filter(.data[[g]] == groups_tbl[[g]][i])
}


v <- gslice[[value_var]]
w <- gslice[[weight_var]]

if (!length(v)) {
  out[[i]] <- tibble(x = numeric(0), cdf = numeric(0))
  next
}

w_total  <- sum(w)
cdf_vals <- vapply(xs,
                   function(x) sum(w[v <= x]) / w_total,
                   numeric(1))

out[[i]] <- bind_cols(
  tibble(x = xs, cdf = cdf_vals),
  as_tibble(groups_tbl[i, , drop = FALSE])
)


}

bind_rows(out)
}

wtname <- wt_final[1]

cdf_perf    <- weighted_cdf_df(dat, "perf_z",     wtname, group_vars = "gender_lab")
cdf_use     <- weighted_cdf_df(dat, "use_best_z", wtname, group_vars = "gender_lab")
cdf_gap     <- weighted_cdf_df(dat, "gap_z",      wtname, group_vars = "gender_lab")
cdf_gap_occ <- weighted_cdf_df(dat, "gap_z",      wtname,
group_vars = c("gender_lab","occ_lab"))

p_perf <- ggplot(cdf_perf, aes(x, cdf, color = gender_lab)) +
geom_line(linewidth = 1) +
labs(
title = "Weighted CDF — Numeracy Performance (z)",
x = "Performance (z)", y = "Cumulative share", color = "Gender"
) +
theme(legend.position = "bottom")

p_use <- ggplot(cdf_use, aes(x, cdf, color = gender_lab)) +
geom_line(linewidth = 1) +
labs(
title = "Weighted CDF — Numeracy Use (z)",
x = "Use (z)", y = "Cumulative share", color = "Gender"
) +
theme(legend.position = "bottom")

p_gap <- ggplot(cdf_gap, aes(x, cdf, color = gender_lab)) +
geom_line(linewidth = 1) +
labs(
title = "Weighted CDF — Confidence–Performance Gap",
x = "Gap (z)", y = "Cumulative share", color = "Gender"
) +
theme(legend.position = "bottom")

p_gap_occ <- ggplot(cdf_gap_occ, aes(x, cdf, color = gender_lab)) +
geom_line(linewidth = 1) +
facet_wrap(~ occ_lab) +
labs(
title = "Weighted CDF — Gap by Occupation Type",
x = "Gap (z)", y = "Cumulative share", color = "Gender"
) +
theme(legend.position = "bottom")

print(p_perf); print(p_use); print(p_gap); print(p_gap_occ)

ggsave("fig_cdf_performance_weighted.png", p_perf,
width = 7, height = 5, dpi = 300)
ggsave("fig_cdf_use_weighted.png",        p_use,
width = 7, height = 5, dpi = 300)
ggsave("fig_cdf_gap_weighted.png",        p_gap,
width = 7, height = 5, dpi = 300)
ggsave("fig_cdf_gap_by_occ_weighted.png", p_gap_occ,
width = 7, height = 5, dpi = 300)

```

##2.4 Self-promotion gap by education and age

```{r}
df_educ <- dat %>%
filter(!is.na(gap_z),
!is.na(educ3),
!is.na(gender_lab))

newdata_gap_educ <- datagrid(
model  = m_gap_educ,
gender = c(0, 1),
educ3  = levels(dat$educ3)
)

pred_gap_educ <- predictions(m_gap_educ, newdata = newdata_gap_educ) %>%
mutate(
gender_text = if_else(gender == 1, "Women", "Men")
)

p_gap_educ <- ggplot() +
geom_jitter(
data = df_educ,
aes(x = educ3, y = gap_z, colour = gender_lab),
alpha = 0.20, width = 0.1, size = 1.1
) +
geom_line(
data = pred_gap_educ,
aes(x = educ3, y = estimate,
colour = gender_text, group = gender_text),
linewidth = 1.2
) +
geom_hline(yintercept = 0, linetype = "dashed") +
labs(
title    = "Self-Promotion Gap by Education Level",
subtitle = "Outcome = Numeracy use z − performance z (gap_z)",
x        = "Education Level (3 groups)",
y        = "Gap (use − perf, z)",
colour   = "Gender"
) +
theme(
axis.text.x      = element_text(angle = 10, hjust = 0.5),
legend.position  = "bottom"
)

print(p_gap_educ)
ggsave("fig_gap_by_education3_gender.png",
p_gap_educ, width = 8, height = 5, dpi = 300)

```

```{r}
newdata_gap_age <- datagrid(
model     = m_gap_age,
gender    = c(0, 1),
age_group = levels(dat$age_group)
)

pred_gap_age <- predictions(m_gap_age, newdata = newdata_gap_age) %>%
mutate(
gender_text = if_else(gender == 1, "Women", "Men")
)

p_gap_age_scatter <- ggplot() +
geom_jitter(
data = dat %>%
filter(!is.na(age_group), !is.na(gap_z), !is.na(gender)),
aes(x = age_group, y = gap_z, colour = gender_lab),
width = 0.15, height = 0, alpha = 0.15, size = 1
) +
geom_hline(yintercept = 0, linetype = "dashed") +
geom_line(
data = pred_gap_age,
aes(x = age_group, y = estimate,
group = gender_text, colour = gender_text),
linewidth = 1.1
) +
labs(
title    = "Self-Promotion Gap by Age Group",
subtitle = "Outcome = Numeracy use z − performance z (gap_z)",
x        = "Age group (AGEG10LFS)",
y        = "Gap (use − perf, z)",
colour   = "Gender"
) +
theme(
axis.text.x     = element_text(angle = 0, hjust = 0.5),
legend.position = "bottom"
)

print(p_gap_age_scatter)
ggsave("fig_gap_by_age_scatter_lines.png",
p_gap_age_scatter, width = 8, height = 5, dpi = 300)

saveRDS(
list(
SVY = list(
m1_gap     = m1_gap,
m_gap_educ = m_gap_educ,
m_gap_age  = m_gap_age
),
meta = list(
objective = "Gender gap in self-promotion (gap_z as outcome); PIAAC Canada; BRR weighted"
)
),
"models_piacc_gender_gap_brr.rds"
)

```

#3. Extension 2: Validating PIAAC Canada Results Using Replication Data

In this extension we test whether patterns from PIAAC Canada align with the replication experiment using linear models on main.dta.

##3.1 Multiple linear regression models with different specifications

```{r}
set.seed(123)

df <- haven::read_dta("data/main.dta")

df <- df %>%
mutate(
female = factor(female),
age_z  = as.numeric(scale(age)),
educ_z = as.numeric(scale(educ_num))
) %>%
filter(!is.na(t_selfpro))

m1 <- lm(t_selfpro ~ female, data = df)
m2 <- lm(t_selfpro ~ female + educ_z, data = df)
m3 <- lm(t_selfpro ~ female + age_z, data = df)
m4 <- lm(t_selfpro ~ female + educ_z + age_z, data = df)

cat("\n=== Multiple Linear Regression Models with Different Specifications  ===\n")
cat("Model 1 (Gender only):      Adj R² =", round(summary(m1)$adj.r.squared, 4), "\n")
cat("Model 2 (+Education):       Adj R² =", round(summary(m2)$adj.r.squared, 4), "\n")
cat("Model 3 (+Age):             Adj R² =", round(summary(m3)$adj.r.squared, 4), "\n")
cat("Model 4 (+Education + Age): Adj R² =", round(summary(m4)$adj.r.squared, 4), "\n")

coef_gender1 <- summary(m1)$coefficients["female1", 1]
coef_gender2 <- summary(m2)$coefficients["female1", 1]
coef_gender3 <- summary(m3)$coefficients["female1", 1]
coef_gender4 <- summary(m4)$coefficients["female1", 1]

cat("\nFemale coefficient change:\n")
cat("  Model 1:", round(coef_gender1, 3), "\n")
cat("  Model 2:", round(coef_gender2, 3), "\n")
cat("  Model 3:", round(coef_gender3, 3), "\n")
cat("  Model 4:", round(coef_gender4, 3), "\n")

```

##3.2 Coefficient comparison table across models
```{r}
get_coef <- function(model, model_name){
broom::tidy(model) %>%
dplyr::mutate(model = model_name) %>%
dplyr::select(model, term, estimate, std.error)
}

coef_df <- dplyr::bind_rows(
get_coef(m1, "Model (1)"),
get_coef(m2, "Model (2)"),
get_coef(m3, "Model (3)"),
get_coef(m4, "Model (4)")
)

coef_df <- coef_df %>%
mutate(term = case_when(
term == "female1"     ~ "β1",
term == "educ_z"      ~ "β2",
term == "age_z"       ~ "β3",
term == "(Intercept)" ~ "β0",
TRUE ~ term
)) %>%
filter(term %in% c("β1","β2","β3","β0"))

coef_df <- coef_df %>%
dplyr::mutate(value = sprintf("%.3f\n(%.3f)", estimate, std.error)) %>%
dplyr::select(term, model, value)

terms_order <- c("β1","β2","β3","β0")

table_wide <- coef_df %>%
mutate(term = factor(term, levels = terms_order)) %>%
arrange(term) %>%
tidyr::pivot_wider(names_from = model, values_from = value)

adj_r2 <- tibble(
term        = "Adj R²",
`Model (1)` = sprintf("%.4f", summary(m1)$adj.r.squared),
`Model (2)` = sprintf("%.4f", summary(m2)$adj.r.squared),
`Model (3)` = sprintf("%.4f", summary(m3)$adj.r.squared),
`Model (4)` = sprintf("%.4f", summary(m4)$adj.r.squared)
)

final_table <- bind_rows(table_wide, adj_r2)
names(final_table)[1] <- ""
final_table[is.na(final_table)] <- ""

final_table[is.na(final_table)] <- ""
names(final_table)[1] <- ""

knitr::kable(
  final_table,
  caption = "Table: Coefficient comparison across four models",
  align   = "lcccc"
) %>%
  kableExtra::kable_styling(
    full_width = FALSE,
    position   = "center",
    font_size  = 11,
    bootstrap_options = c("condensed")
  )


```

##3.3 Gender gap ~ education (continuous): interaction model
```{r}
set.seed(123)

df <- df %>%
mutate(
female = ifelse(female == 1, "Female", "Male"),
female = factor(female, levels = c("Male", "Female"))
)

if(!("educ_z" %in% names(df))) {
df <- df %>%
mutate(educ_z = scale(educ_num)[,1])
}

model_educ <- lm(t_selfpro ~ female * educ_z, data = df)

pred_df <- expand.grid(
educ_z = seq(min(df$educ_z, na.rm = TRUE),
max(df$educ_z, na.rm = TRUE),
length.out = 100),
female = c("Male", "Female")
)

pred_df$female <- factor(pred_df$female, levels = c("Male","Female"))
pred_df$pred   <- predict(model_educ, newdata = pred_df)

```

```{r}
set.seed(123)

p <- ggplot(pred_df, aes(x = educ_z, y = pred, color = female)) +
geom_line(size = 1.4) +
scale_color_manual(
values = c("Male" = "#2C73B0", "Female" = "#E05252"),
name   = "Gender"
) +
labs(
title    = "Belief–Performance Gap Across Education (continuous)",
subtitle = "Model: t_selfpro ~ female * educ_z",
x        = "Education (z-scored)",
y        = "Belief − Performance\n(Self-predicted minus actual score)"
) +
theme_minimal(base_size = 16) +
theme(
plot.title    = element_text(face = "bold", size = 20),
plot.subtitle = element_text(size = 14),
axis.title.y  = element_text(vjust = 1.2)
)

p

ggsave(
filename = "belief_performance_gap_by_education.png",
plot     = p,
width    = 10,
height   = 6,
dpi      = 300
)

```

#Appendix: Extended Methodological Notes for Extension 2 
(How Education and Age Shape the Gender Confidence Gap) 


```{r}
set.seed(123)

df <- haven::read_dta("data/main.dta") %>%
mutate(
female   = as.integer(female),
female_f = factor(female, levels = c(0,1), labels = c("Male","Female"))
) %>%
filter(!is.na(t_selfpro), !is.na(female), !is.na(educ_num))

edu_labels <- c(
"1 = No schooling / Primary",
"2 = Some secondary",
"3 = Completed secondary / High school",
"4 = Some college / Associate degree",
"5 = Bachelor’s degree",
"6 = Some graduate study",
"7 = Master’s degree",
"8 = Professional degree",
"9 = Doctorate (PhD)"
)
cat("\nEducation code meanings:\n")
cat(paste(edu_labels, collapse = "\n"))

by_code <- df %>%
group_by(educ_num, female) %>%
summarise(mean_self = mean(t_selfpro, na.rm=TRUE), n = n(), .groups="drop") %>%
pivot_wider(names_from = female, values_from = c(mean_self,n), names_sep="_") %>%
mutate(gap_FminusM = mean_self_1 - mean_self_0,
N_total = n_0 + n_1)
print(by_code)

scan_tbl <- lapply(1:8, function(cut){
tmp <- df %>% mutate(group = ifelse(educ_num <= cut, "LowEdu", "HighEdu"))
mL  <- lm(t_selfpro ~ female, data = filter(tmp, group=="LowEdu"))
mH  <- lm(t_selfpro ~ female, data = filter(tmp, group=="HighEdu"))
tibble(
cutpoint = cut,
gap_low  = coef(mL)["female"],
gap_high = coef(mH)["female"],
diff_gap = gap_high - gap_low,
n_low    = nrow(filter(tmp, group=="LowEdu")),
n_high   = nrow(filter(tmp, group=="HighEdu"))
)
}) %>% bind_rows()

cat("\n=== Difference in gender gap (High − Low) across cutpoints ===\n")
print(scan_tbl)

ggplot(scan_tbl, aes(x = cutpoint, y = diff_gap)) +
geom_hline(yintercept = 0, linetype="dashed", color="gray40") +
geom_line(color="steelblue", linewidth=1) +
geom_point(size=2, color="orange") +
scale_x_continuous(
breaks = 1:8,
labels = c(
"1|2\n(No schooling\n→ Secondary)",
"2|3\n(Secondary\n→ High School)",
"3|4\n(High School\n→ College)",
"4|5\n(College\n→ Bachelor)",
"5|6\n(Bachelor\n→ Grad)",
"6|7\n(Grad\n→ Master)",
"7|8\n(Master\n→ Professional)",
"8|9\n(Professional\n→ PhD)"
)
) +
labs(
title = "Change in Gender Gap (High − Low) Across Education Splits",
subtitle = "Positive values → larger gap in higher-education group",
x = "Cutpoint (education code ≤ cut = Low group)",
y = "Δ Gender Gap (High − Low)"
) +
theme_minimal(base_size = 14) +
theme(
plot.title = element_text(face = "bold", size = 22, hjust = 0.5),
plot.subtitle = element_text(size = 15, hjust = 0.5),
axis.title = element_text(size = 14),
axis.text.x = element_text(size = 11),
axis.text.y = element_text(size = 11)
)

ggsave(
filename = "gender_gap_sensitivity.png",
width = 10,
height = 6,
dpi = 300
)

```
## A2. Heterogeneity in the Gender Confidence Gap Across Performance Levels

## A2.1 Compute bin-specific gender gaps

```{r heterogeneity-bins}
# Use the same regression sample (df_t2a) from Table 2 Panel A analysis
# Compute belief_gap = belief - performance for heterogeneity analysis

# First, we need to get the actual belief variable (in 0-20 scale, not 0-100)
# The original data has 'belief' which is the absolute number correct believed

# Re-extract belief from main data matching the filtered sample
belief_col_name <- pick_col(main, c("belief", "predicted_correct", "belief_num"))

df_hetero_base <- main %>%
  mutate(
    female_id = as.numeric(.data[[female_col]]),
    num_correct_id = as.numeric(.data[[perf_col]]),
    belief_absolute = as.numeric(.data[[belief_col_name]])  # 0-20 scale
  )

# Apply same filters as main analysis
if (best$treat == "on" && length(treat_col) == 1) {
  vals <- unique(na.omit(df_hetero_base[[treat_col]]))
  if (all(vals %in% c(0, 1, 0L, 1L))) {
    df_hetero_base <- filter(df_hetero_base, .data[[treat_col]] == 1)
  } else {
    df_hetero_base <- filter(df_hetero_base, is_trueish(.data[[treat_col]]))
  }
}

if (!is.na(wave_col) && best$wave != "none") {
  df_hetero_base <- filter(df_hetero_base, .data[[wave_col]] == 
                            suppressWarnings(type.convert(best$wave, as.is = TRUE)))
}

if (best$eval == "on" && length(alleval_col) == 1) {
  df_hetero_base <- filter(df_hetero_base, .data[[alleval_col]] %in% c(1, "1", TRUE))
}

if (best$math == "on" && length(allmath_col) == 1) {
  df_hetero_base <- filter(df_hetero_base, .data[[allmath_col]] %in% c(1, "1", TRUE))
}

df_hetero <- df_hetero_base %>%
  mutate(
    belief_gap = belief_absolute - num_correct_id,  # Both on 0-20 scale
    gender = if_else(female_id == 1, "Female", "Male"),
    female = female_id,
    num_correct = num_correct_id
  ) %>%
  select(gender, female, num_correct, belief_gap) %>%
  drop_na()

# Define performance bins: Low (0-6), Medium (7-13), High (14-20)
df_hetero <- df_hetero %>%
  mutate(
    perf_bin = cut(
      num_correct,
      breaks = c(-0.5, 6.5, 13.5, 20.5),
      labels = c("Low (0–6)", "Medium (7–13)", "High (14–20)")
    )
  )

# Compute bin-specific gender means and gender gap (Male - Female)
bin_stats <- df_hetero %>%
  group_by(perf_bin, gender) %>%
  summarise(
    n = n(),
    mean_gap = mean(belief_gap, na.rm = TRUE),
    var_gap = var(belief_gap, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_wider(
    names_from = gender,
    values_from = c(n, mean_gap, var_gap)
  ) %>%
  mutate(
    # Gender gap: Male mean - Female mean
    delta = mean_gap_Male - mean_gap_Female,
    # Standard error of the difference
    se_delta = sqrt(var_gap_Male / n_Male + var_gap_Female / n_Female),
    # 95% confidence interval
    ci_lo = delta - 1.96 * se_delta,
    ci_hi = delta + 1.96 * se_delta
  ) %>%
  mutate(
    perf_bin = factor(
      perf_bin,
      levels = c("Low (0–6)", "Medium (7–13)", "High (14–20)")
    )
  ) %>%
  arrange(perf_bin)

# Display results
print(bin_stats)

# Save table
write_csv(bin_stats, "belief_gap_gender_diff_by_perf_bin.csv")
```

## A2.2 Visualize ECDFs by performance band

```{r heterogeneity-ecdf, fig.width=12, fig.height=4.5}
# Create ECDF plots faceted by performance band
# This shows how the gender gap in belief-performance varies across ability levels

p_hetero <- df_hetero %>%
  mutate(gender = factor(gender, levels = c("Female", "Male"))) %>%
  ggplot(aes(x = belief_gap, color = gender, linetype = gender)) +
  stat_ecdf(linewidth = 1.1) +
  facet_wrap(~ perf_bin, ncol = 3, scales = "free_x") +
  geom_vline(xintercept = 0, color = "red", alpha = 0.6, linewidth = 0.5) +
  scale_color_manual(values = c("Female" = "black", "Male" = "blue")) +
  scale_linetype_manual(values = c("Female" = "solid", "Male" = "dashed")) +
  theme_classic(base_size = 12) +
  theme(
    legend.position = "bottom",
    strip.background = element_rect(fill = "grey90", color = "black"),
    strip.text = element_text(face = "bold", size = 11)
  ) +
  labs(
    title = "Heterogeneity Analysis: Belief – Performance (ECDF) by Performance Band",
    x = "Belief – Performance",
    y = "CDF",
    color = "Gender",
    linetype = "Gender"
  )

print(p_hetero)

# Save figure
ggsave("figure_ecdf_belief_minus_performance_by_perf_bin.pdf",
       p_hetero, width = 12, height = 4.5)
ggsave("figure_ecdf_belief_minus_performance_by_perf_bin.png",
       p_hetero, width = 12, height = 4.5, dpi = 300)
```
